---
  title: "Art of Statistics: Table 5.1 (page 123) and Figure 5.1 (page 124) Galton family heght data"
  #output: md_document
  output: html_document
---

 Data from 1991-1995 are contained in [05-1-galton-x.csv](05-1-galton-x.csv),
Although the book says the data is from [HistData: Data Sets from the History of Statistics and Data Visualization, 2018](https://cran.r-project.org/web/packages/HistData/index.html), I have actually used this version of  [Galton's Height Data](http://www.randomservices.org/random/data/Galton.html)

```{r}


galton<-read.csv("05-1-galton-x.csv",header=TRUE) # read csv file into dataframe galton
attach(galton) #uncomment if/while necessary

summary(galton)

# summary statistics
# need means for unique fathers and mothers - identify first mention of each family
Unique.Fathers=numeric()
Unique.Mothers=numeric()
nunique=1 # number of unique families 
Unique.Fathers[1] = Father[1]
Unique.Mothers[1] = Mother[1]  
 for(i in 2:length(Family))
{   
    if(Family[i] != Family[i-1]){
      nunique=nunique+1
    Unique.Fathers[nunique]=Father[i]
    Unique.Mothers[nunique]=Mother[i]
    }
  }
  
length(Unique.Fathers)
summary(Unique.Fathers)
sd(Unique.Fathers)

length(Unique.Mothers)
summary(Unique.Mothers)
sd(Unique.Mothers)
 
Son = Height[Gender=="M"] 
length(Son)
summary(Son)
sd(Son)

Daughter = Height[Gender=="F"]
length(Daughter)
summary(Daughter)
sd(Daughter)
```


### Figure 5.1 (page 124) Linear regression of sons' on fathers' heights

```{r}

# Heights of fathers of sons
FatherS = Father[Gender=="M"]

fit <- lm(Son ~ FatherS) # linear regression data in fit
Predicted <- predict(fit)   # Get the predicted values
summary(fit)

FatherS.j <- jitter(FatherS, factor=5) 
Son.j <- jitter(Son, factor=5)

xlims=ylims=c(55,80)
par(mfrow=c(1,1), mar=c(4,4,2,0), pty="s")  # square plot

plot(FatherS.j, Son.j, xlim=xlims,ylim=ylims,cex=0.7,
     xlab="father's height (inches)",ylab="son's height (inches)" , col="gray68")
lines(c(xlims[1],xlims[2]),c(xlims[1],xlims[2]),lty=2 )
lines(Predicted~FatherS,lwd=2)

```

### Now in ggplot

```{r}
library(ggplot2)
# create new data frame with exact and jittered, and predcted values

Males = cbind.data.frame(FatherS,FatherS.j,Son,Son.j,Predicted)

p <- ggplot(Males, aes(x=FatherS, y=Son)) # initial plot object
p <- p + geom_point(x=FatherS.j,y=Son.j,shape= 1) # defines scatter type plot
p <- p + labs(x="Father's height (inches)", y= "Son's height (inches)") # adds x and y axis labels
p <- p + theme(legend.position="none")#, legend.box = "horizontal") # removes the legend
p <- p + expand_limits(x = c(55,80),y = c(55,80)) # expand the axis limits
p <- p + geom_line(aes(FatherS,Predicted),size=1.5) # add previously fitted linear regression line  

p <- p + geom_abline(slope=1, linetype="dashed") # line to represent equality between son and father height


# select single data points by CSV datarow numbers
pointA=c(137)  
pointB=c(28)

# plot residual line and end points for selectedpointA
p <- p + geom_point(aes(x=FatherS.j[pointA], y = Predicted[pointA]), shape = 1)
p <- p + geom_point(aes(x=FatherS.j[pointA], y = Son.j[pointA]), shape = 1)
p <- p + geom_segment(linetype="dashed", size=1, colour="purple",aes(x=FatherS.j[pointA],y=Son.j[pointA],xend = FatherS.j[pointA], yend = Predicted[pointA])) #p <- p + p

# plot residual line and end points for pointB
p <- p + geom_point(aes(x=FatherS.j[pointB], y = Predicted[pointB]), shape = 1)
p <- p + geom_point(aes(x=FatherS.j[pointB], y = Son.j[pointB]), shape = 1)
p <- p + geom_segment(linetype="dashed", size=1, colour="purple",aes(x=FatherS.j[pointB],y=Son.j[pointB],xend = FatherS.j[pointB], yend = Predicted[pointB]))

p #displays the result

```

Figure 5.1 Scatter of heights of 465 fathers and sons from Galton's data (many fathers are repeated since they have multiple sons). A jitter has been added to separate the points, and the diagonal dashed line represents exact equality between son and father's heights. The solid line is the standard 'best-fit' line. Each point gives rise to a 'residual' (dashed line), which is the size of the error were we to use the line to predict a son's height from his father's.