---
  title: "Art of Statistics: Table 2.1 (page 50), Figure 2.2 (page 42), Figure 2.3 (page 45)  Jelly beans counts"
  #output: md_document
  output: html_document
---
### Table 2.1: Summary statistics for Jelly bean counts 

Data comprise 915 guesses of the  numnber of jelly-beans in a jar, and are contained in [02-1-bean-data-full-x.csv](02-1-bean-data-full-x.csv).

 ```{r}
Guesses<-read.csv("02-1-bean-data-full-x.csv",header=FALSE) # read in data
NGuesses <- nrow(Guesses)
summary(Guesses)
sd(Guesses$V1) # standard deviation of first column of data-frame
```


### Figure 2.2: Different displays of Jelly bean counts 

 ```{r}
library(magrittr)
library("ggplot2")
library("ggpubr")

# load ggplot2 themes into BlankXTheme, BlankYTheme, and BiggerTicks

BlankXTheme <- theme(axis.title.x=element_blank(),
                     axis.text.x=element_blank(),
                     axis.ticks.x=element_blank())

BlankYTheme <- theme(axis.title.y=element_blank(),
                     axis.text.y=element_blank(),
                     axis.ticks.y=element_blank(),
                     plot.margin=unit(c(0.1,0.1,0.1,0.65),"cm")) # also give a little space for plot labelling

BiggerTicks <- theme(axis.text.x = element_text(size=16), axis.title.x = element_text(size=16))

AugmentedGuesses <- cbind(Guesses, runif(NGuesses), rep(1, NGuesses))
names(AugmentedGuesses) <- c("Guess", "Aug", "Group") # Aug is jittered plotting position

Strip <- ggplot(AugmentedGuesses, aes(x=Guess, y=Aug)) + geom_point() + labs(y="") + theme_bw() +
          BlankXTheme + BlankYTheme # scatter type plot

Box <- ggplot(AugmentedGuesses, aes(x=Group, y=Guess)) + geom_boxplot() + coord_flip() + theme_bw() + 
          BlankXTheme + BlankYTheme # box and whisker type plot

Hist<- ggplot(AugmentedGuesses, aes(x=Guess)) + geom_histogram(bins=50) + theme_bw()  + BlankYTheme + BiggerTicks + labs(x="Guess at number of beans in jar", hjust=0.0) # histogram plot

p <- ggarrange(Strip, Box, Hist, ncol=1, nrow=3, 
          labels=c("(a)","(b)","(c)"), hjust=0.0) # arrange in grid with a,b,c labels for the plots, 
p
```

Figure 2.2 Different ways of showing the pattern of 915 guesses of the number of jelly beans in the jar. (a) A strip-chart or dot-diagram, with a jitter to prevent points lying on top of each other;  (b) a box-and-whisker plot; (c) a histogram

### Figure 2.3: Different displays of Jelly bean counts - logarithmic scale

 ```{r}

logbreaks <- c(200,500,2000,5000,20000)

Strip <- ggplot(AugmentedGuesses, aes(x=Guess, y=Aug)) + geom_point() + 
    scale_x_continuous(trans="log10", breaks=logbreaks) + 
    theme_bw() + BlankXTheme + BlankYTheme # scatter type plot

Box <- ggplot(AugmentedGuesses, aes(x=Group, y=Guess)) + geom_boxplot() + 
  scale_y_continuous(trans="log10", breaks=logbreaks) + 
  theme_bw() + BlankXTheme + BlankYTheme + coord_flip() # box and whisker type plot

Hist<- ggplot(AugmentedGuesses, aes(x=Guess)) + geom_histogram(bins=50) + 
  scale_x_continuous(trans="log10", breaks=logbreaks) + 
  theme_bw() + BlankYTheme + BiggerTicks + labs(x="Guess at number of beans in jar") # histogram plot

p <- ggarrange(Strip, Box, Hist, ncol=1, nrow=3, 
          labels=c("(a)","(b)","(c)"), hjust=0.0) # arrange plots in grid with labels
p
```

Figure 2.3 Graphical displays of the jelly-bean guesses plotted on a logarithmic scale. (a) Strip-chart; (b) box-and-whisker plot; (c) histogram all show a fairly symmetric pattern.